---
title: "Lab 2 (D Glave"
author: "Dylan Glave"
date: "1/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(palmerpenguins)
library(here)
library(lubridate)
# for PCA
library(ggfortify)

# for ggplot customization
library(readxl)
library(gghighlight)
library(plotly)

# for compound figures
library(patchwork)
```


## Part 1: PCA with Penguins

```{r}
# going to scale, and breakdown by PCs

penguins_pca <- penguins %>% 
  select(body_mass_g, ends_with("_mm")) %>% 
  drop_na() %>% 
  scale() %>% 
  prcomp()

penguins_pca$rotation
#Above, rotation shows the loadings (vector length) of each variable with each PC
#Below, autoplot automatically selects the type of plot that I want to make in ggplot. Since we've converted penguins_pca into a list for of PCs, it assumed we wnat a PCA biplot
# Making a df of penguin info that matches the PCA. we'll have the same observation set, but will ahve info like species.

penguin_complete <- penguins %>% 
  drop_na(body_mass_g, ends_with("_mm"))
# we use this MATCHING data to add aesthetics to the biplot
autoplot(penguins_pca, 
         data = penguin_complete,
         colour = 'species',
         loadings = TRUE,
         loadings.label = TRUE) +
  theme_minimal()
# All the customization in ggplot is available in autoplot
# facto_extra is another great place to play with biplots/PCA 
```

## Part 2: New skills for ggplot customization + Reading in different file types

### Starting with .xlsx file

```{r}
fish_noaa <- read_excel(here("data", "foss_landings.xlsx")) %>% 
  clean_names() %>% 
  mutate(across(where(is.character), tolower)) %>% 
  mutate(fish_name= str_sub(nmfs_name, end = -4)) %>% 
  filter(confidentiality == "public")
# Instead of adding columns, we're transforming columns w that function.for any character column, we send it to lower case. Across selects different columns, we chose all character columns (could be "ends_with("mm"). and within that we did 
# We then used string_sub within mutate (stringr) to remove the last 4 characters.
```

Make a customized graph: 

```{r}
fish_plot <- ggplot(data =  fish_noaa, 
                    aes( x = year, y = pounds)) + 
  geom_line(aes(color = fish_name), show.legend = FALSE) +
  theme_minimal()

fish_plot

ggplotly(fish_plot)
#this makes any map interactive

### use gghighlight to highlight certain series. 

fish_plot_2 <- ggplot(data =  fish_noaa, 
                    aes( x = year, y = pounds, group = fish_name)) +
  geom_line() +
  theme_minimal() +
  gghighlight(fish_name == "tunas") + # Highlight just tunas
  theme_minimal()

fish_plot_2

fish_plot_3 <- ggplot(data =  fish_noaa, 
                    aes( x = year, y = pounds, group = fish_name)) +
  geom_line() + # If we add aes to this, it only applies to gghighlighted area
  theme_minimal() + 
  gghighlight(max(pounds) > 1e8) + 
    ###Highlight just series that have weights  over 10^8
  theme_minimal()

fish_plot_3
```
 We now know to how to make interactive graphs, AND highlight things! 
 
 
####  Read in data from a URL, lubridate() refresher, mutate() as a workhorse to transform variables


```{r}
monroe_wt <- read_csv("https://data.bloomington.in.gov/dataset/2c81cfe3-62c2-46ed-8fcf-83c1880301d1/resource/13c8f7aa-af51-4008-80a9-56415c7c931e/download/mwtpdailyelectricitybclear.csv") %>% 
  clean_names()

# Let's make a graph BY MONTH of how much Kwh is used
#Start by making dates a date. 
#Make a new column to add month names to the numbered month
# And then we'll change it to a factor, based on the number of months


```


```{r}
monroe_ts <- monroe_wt %>% 
  mutate(date = mdy(date)) %>% # Convert date to a stored date
  mutate(record_month = month(date)) %>% 
  mutate(month_name = month.abb[record_month]) %>% # This switches all recorded months to month names, as month.abb
mutate(month_name = fct_reorder(month_name, record_month)) #change from character to factor, AND set the order based on another variable

  
  ggplot(data = monroe_ts, aes(x = month_name, y = total_kWh)) +
  geom_jitter(aes(color = month_name),
              show.legend = FALSE,
              alpha = 0.7, 
              size = 0.3,
              width = 0.4)
```

### Part 3: Compound figures with 'patwork' 

```{r}
graph_a <- ggplot(data = penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point()

graph_b <- ggplot(data = penguins, aes(x = species, y = flipper_length_mm)) +
  geom_jitter(aes(color = species))

# Use | vertical line| side by side
# use / to put graphs on top of one another

graph_c <-  graph_a | graph_b #They go side by side! 
(graph_a | graph_b) / fish_plot_3 & theme_dark()    
#Use pemdas t put more figures together, via patchwork. 

graph_c

ggsave(here("fig", "graph_c_dg.png")) # you can customize width and height of figures, by specifying them within ggsave. 
```


## Part 2: Creating a Shiny App. 

When we make an App, it's creates a new, moveable folder. 


